Visually pleasing knot projections

### 2023-02-02 inst/jma_rebuttal_2023_01_30.tex

\section*{Detailed rebuttal}

Comments from the Editors and Reviewers:

I appreciate and accept most of the comments made by the author in his
rebuttal and the corresponding fixes in the manuscript.  But I have a
remaining concern that much of this material is perfect for an on-line
tutorial on how to use package [27] to draw optimized knot diagrams,
but JMA is not the right place to publish such a tutorial.

The content that is most fitting for a JMA paper is a discussion of
what makes a knot diagram ``beautiful" and that discussion will
clearly be biased by the personal view of the author (perhaps enhanced
by a ``user-study" among family and colleagues).  Thus lines 14 - 20
on page 4 are of prime importance and could be fleshed out more.

@@

The main contribution to JMA would then be a detailed discussion of
how the author has struggled to capture these goals in the form of
mathematical functions that can be optimized and will then approximate
these goals.  In this respect, the manuscript still has some
shortcomings.

They may be fixed in two ways: Enhance the detailed explanations of
the choices made in creating package [27]; and add a more extended
discussion of a few things that might be approached differently in a
future package ( -- which can also handle links).

In view of this, I recommend that Appendices A and B should be firmly
integrated into the main text.  The discussions therein relate rather
directly to what I see as the main contribution of a JMA paper on this
topic.\textcolor{blue}{\\Done.\\ \\}

### 2023-02-03 inst/jma_rebuttal_2023_01_30.tex

For each crossing point, the user would see an orthogonal ``cross-hair"
that can be shifted and rotated as a whole.  In addition, there are
two parameters (handle-pairs) that control the ``velocities" of the two
strands through the crossing.  The optimization routine would
fine-tune these velocities to achieve matching curvature for the two
sequentially connected Bezier curves.

And, yes, if a crossing point falls on a line of mirror symmetry, the
crosshair would automatically be angled at 45 degrees with respect to
this line.  And if the knot has n-fold rotational symmetry, n such
cross hairs would be placed at the vertices of a regular n-gon.  (All
this is not such a ``nightmare" if it is properly considered from the
beginning.)

I don't expect that anything in the current system should be changed;
but this approach should definitely be considered in a future version
of such a package.
\textcolor{blue}{\\ This scheme certainly sounds achievable and I have
  alluded to it in the manuscript at the two trefoil knot
  discussion.\\ \\}

RE: Mismatched diagram and symmetry specifications: ``This situation
cannot occur" -- If the user makes a mistake, will the current system
send a suitable error message?\textcolor{blue}{\\For convenience, I
  reproduce the original exchange below:\\ \\} {\tiny Reviewer's
  comment, round 3: ``If all nodes fulfill the specified symmetry
  constraints (as shown on page 11, lines 15-22 and on lines 35-43),
  but the connection diagram does not completely adhere to the
  specified symmetry - What happens? - Does the package check this and
  let the user know about the discrepancy?"}  \textcolor{blue}{\tiny
  My original reply: ``This situation cannot occur: symmetric nodes
  imply a symmetric diagram".\\ \\} \textcolor{blue}{A more nuanced
  reply would have been for me to observe that in the event of a
  mismatch between diagram and symmetry specifications, the symmetry
  specifications take priority and the resulting symmetrised diagram
  automatically accommodates the imposed symmetry.  In practice what
  happens is immediately visually apparent: one sees beautiful
  symmetric nodes joined by Bezier curves that do not possess the
  intended topology.  The diagram is a disorganised and ugly tangled
  mess, clearly not reflecting the intended configuration.  In
  practice, one glance at the diagram simultaneously reveals the
  problem and its solution (just reconnect the dots in the correct
  order).\\ \\}

### 2023-02-05 inst/jma_rebuttal_2023_01_30.tex

\textcolor{blue}{ Short story: }

\section*{Detailed rebuttal}

Comments from the Editors and Reviewers:

I appreciate and accept most of the comments made by the author in his
rebuttal and the corresponding fixes in the manuscript.  But I have a
remaining concern that much of this material is perfect for an on-line
tutorial on how to use package [27] to draw optimized knot diagrams,
but JMA is not the right place to publish such a
tutorial.\textcolor{blue}{I have expanded the discussion of the
  aesthetics of knot diagrams, but would point out that capturing the
  goals}

@@

The content that is most fitting for a JMA paper is a discussion of
what makes a knot diagram ``beautiful" and that discussion will
clearly be biased by the personal view of the author (perhaps enhanced
by a ``user-study" among family and colleagues).  Thus lines 14 - 20
on page 4 are of prime importance and could be fleshed out more.

The main contribution to JMA would then be a detailed discussion of
how the author has struggled to capture these goals in the form of
mathematical functions that can be optimized and will then approximate
these goals.  In this respect, the manuscript still has some
shortcomings.

They may be fixed in two ways: Enhance the detailed explanations of
the choices made in creating package [27]; and add a more extended
discussion of a few things that might be approached differently in a
future package ( -- which can also handle links).
\textcolor{blue}{\\I have added discussions as suggested and also
  included a modified analysis of functionality for links.\\ \\}

### 2023-02-05 inst/jma_rebuttal_2023_01_30.tex

\documentclass[12pt]{article}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{fancyvrb}
\begin{document}

\section*{Rebuttal to ``Visually pleasing knot projections" by
  Robin K. S. Hankin (submitted to JMA, manuscript 209604076); review
  30 January 2023}

@@

(Below, the reviewer's comments are in black, and my replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate.

\section*{\textcolor{blue}{Overall response to the review}}

\textcolor{blue}{The comments have resulted in a much stronger
  document.  I have carried out all the detailed suggestions, as
  presented below.  My take on the comment reading ``much of this
  material is perfect for an on-line tutorial" requires a more
  subjective and personal response.  I would characterise the
  submission, not as a online tutorial, but as a discursive yet
  mathematical outlining of the conceptual and computational
  difficulties of automated beautification of knot diagrams (not
  forgetting to discuss the advantages!).  Automated beautification of
  knot diagrams sounds easy, and I originally thought it would take me
  a couple of hours.  I cannot be the first person to try to do this;
  much of the submission could be viewed as documenting the
  difficulties of the only known solution to this problem.\\ \\ I do
  not regard the essence of the submission as being at all dependent
  on inkscape or R.  Any vectorised graphics editor (VGE) that can
  work with Bezier curves would be as good as inkscape.  Adobe
  illustrator, or Google drawings etc etc could be substituted for
  inkscape.  So in the submission, ``inkscape'' is short for ``any
  VGE''.  The difficulty is that a manuscript needs a specfic VGE to
  concretify the discussions.  The same reasoning applies to my choice
  of R, which was again driven by my familiarity with it; matlab or
  python would be just as good.  I believe that the resubmission gets
  this point across.}

@@

\section*{Detailed rebuttal}

Comments from the Editors and Reviewers:

I appreciate and accept most of the comments made by the author in his
rebuttal and the corresponding fixes in the manuscript.  But I have a
remaining concern that much of this material is perfect for an on-line
tutorial on how to use package [27] to draw optimized knot diagrams,
but JMA is not the right place to publish such a
tutorial.\textcolor{blue}{\\I have expanded the discussion of the
  aesthetics of knot diagrams, but would point out that capturing the
  goals\\ \\}

### 2023-02-06 inst/jma_rebuttal_2023_01_30.tex

\documentclass[12pt]{article}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{fancyvrb}
\begin{document}

\section*{Rebuttal to ``Visually pleasing knot projections" by
  Robin K. S. Hankin (submitted to JMA, manuscript 209604076); review
  30 January 2023}

(Below, the reviewer's comments are in black, and my replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate).

\section*{\textcolor{blue}{Overall response to the review}}

\textcolor{blue}{The comments have resulted in a much stronger
  document.  I have carried out all the detailed suggestions, as
  presented below, and offer a commentary on some of the ideas.}

@@

\section*{Detailed rebuttal}

Comments from the Editors and Reviewers:

I appreciate and accept most of the comments made by the author in his
rebuttal and the corresponding fixes in the manuscript.  But I have a
remaining concern that much of this material is perfect for an on-line
tutorial on how to use package [27] to draw optimized knot diagrams,
but JMA is not the right place to publish such a
tutorial.\textcolor{blue}{ I would characterise the submission, not as
  a online tutorial, but as a discursive yet mathematical outlining of
  the conceptual and computational difficulties of automated
  beautification of knot diagrams (not forgetting to discuss the
  advantages!).  I have expanded the discussion of the aesthetics of
  knot diagrams.  \\ \\ On the online tutorial comparison: I do not
  regard the essence of the submission as being at all dependent on
  inkscape or R.  Any vectorised graphics editor (VGE) that can work
  with Bezier curves would be as good as inkscape.  Adobe illustrator,
  or Google drawings etc etc could be substituted for inkscape.  So in
  the submission, ``inkscape'' is short for ``any VGE''.  The
  difficulty is that a manuscript needs a specific VGE to concretify
  the discussions.  The same reasoning applies to my choice of R,
  which was again driven by my familiarity with it; matlab or python
  would be just as good.  I believe that the resubmission gets this
  point across.}

The content that is most fitting for a JMA paper is a discussion of
what makes a knot diagram ``beautiful" and that discussion will
clearly be biased by the personal view of the author (perhaps enhanced
by a ``user-study" among family and colleagues).  Thus lines 14 - 20
on page 4 are of prime importance and could be fleshed out more.
\textcolor{blue}{\\I have added a some discussion as suggested.}

@@

For each crossing point, the user would see an orthogonal ``cross-hair"
that can be shifted and rotated as a whole.  In addition, there are
two parameters (handle-pairs) that control the ``velocities" of the two
strands through the crossing.  The optimization routine would
fine-tune these velocities to achieve matching curvature for the two
sequentially connected Bezier curves.

And, yes, if a crossing point falls on a line of mirror symmetry, the
crosshair would automatically be angled at 45 degrees with respect to
this line.  And if the knot has n-fold rotational symmetry, n such
cross hairs would be placed at the vertices of a regular n-gon.  (All
this is not such a ``nightmare" if it is properly considered from the
beginning.)

I don't expect that anything in the current system should be changed;
but this approach should definitely be considered in a future version
of such a package.
\textcolor{blue}{\\ This scheme certainly sounds achievable and I have
  alluded to it in the manuscript at the two trefoil knot
  discussion.\\ \\}

RE: Mismatched diagram and symmetry specifications: ``This situation
cannot occur" -- If the user makes a mistake, will the current system
send a suitable error message?\textcolor{blue}{\\For convenience, I
  reproduce the original exchange below:\\ \\} {\tiny Reviewer's
  comment, round 3: ``If all nodes fulfill the specified symmetry
  constraints (as shown on page 11, lines 15-22 and on lines 35-43),
  but the connection diagram does not completely adhere to the
  specified symmetry - What happens? - Does the package check this and
  let the user know about the discrepancy?"}  \textcolor{blue}{\tiny
  My original reply: ``This situation cannot occur: symmetric nodes
  imply a symmetric diagram".\\ \\} \textcolor{blue}{A more nuanced
  reply would have been for me to observe that in the event of a
  mismatch between diagram and symmetry specifications, the symmetry
  specifications take priority and the resulting symmetrised diagram
  has symmetric nodes but the connections are not as intended.  In
  practice, this is immediately visually apparent: one sees beautiful
  symmetric nodes joined by Bezier curves that do not possess the
  intended topology.  The diagram is a disorganised and ugly tangled
  mess, clearly not reflecting the intended configuration.  Generally,
  one glance at the diagram simultaneously reveals the problem and its
  solution (just reconnect the dots in the correct order).\\ \\}

@@

RE: A hierarchical way of dealing with symmetry: ``This is exactly
what I have done!  The referee gives an accurate summary of the code's
methodology." - Place a corresponding, extended description into the
main text.\textcolor{blue}{\\Done.  New text added in section 3.2:
  From a numerical perspective, the software takes advantage of a
  hierarchical representation of symmetrical knots: the curve is
  specified for one unique fragment of the knot, and the complete knot
  is reconstructed by replicating that fragment with mirroring and
  rotation as required.  The optimization is effectively carried out
  for this unique fragment (with proper end-conditions).\\ \\}

RE: A slider interface for changing some optimization parameters: It
is clear why this would not be practical in the current system, where
the optimization of a pretzel knot ``may take a few weeks."  But this
also suggests that there is much room for improvement!  Perhaps some
badness functions are not really needed or are badly implemented.
\textcolor{blue}{\\ This is an excellent suggestion and indeed many of
  the badness functions may well turn out to be superfluous, or at
  least susceptible to faster heuristics [although I would argue that
    the functions all adhere to efficient vectorised R coding
    standards].  If it turns out that some of the badness weightings
  can be set to zero without degrading the resulting visual
  attractiveness, this would certainly be worth pursuing.\\ \\}

RE: Link projections: In a future system, the planning for handling
links should come in right at the beginning.  It is not at all clear
why ``Each component would have its own badness."

@@

Page 6, Line 42: What is ``{\tt > b <}" ?  \textcolor{blue}{\\ This is
  just a detail of standard R syntax.  The first ``{\tt > }" is just
  the R prompt; letter {\tt b} is just a variable; and the symbol
  ``{\tt <-}" usually has a gloss of ``becomes".  Thus {\tt b <- 4} is
  read ``{\tt b} becomes 4". The reason I need a variable is that I
  want to show the result of applying two functions, {\tt
    total\_crossing\_angles()} and {\tt total\_bending\_energy()}, to
  the same variable {\tt b}.\\ \\}

Figures 3 and 5: Try running these figures with only 50 points (and
circles) for each Bezier segment. \textcolor{blue}{\\This is easily
  achievable in the package but the result fails to provide any
  informative or enlightening information.  It simply highlights the
  fact that $n$ was too small.  Using $n=50$ is (almost)
  indistinguishable from $n=100$.  I found that $n=100$ represented
  essentially the best visual representation while being
  computationally not too onerous.  I show the requested figure at the
  end (along with $n=7$ and $n=100$ for comparison),
  figures~\ref{7_6_n7}, \ref{7_6_n50}, and \ref{7_6_n100}.  These were
  created with}

\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{blue}{R> pdf(file="k7_6_n07.pdf") ; knotplot2(k7_6,n=7)   ; dev.off()}
\end{Verbatim}
\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{blue}{R> pdf(file="k7_6_n50.pdf") ; knotplot2(k7_6,n=50)  ; dev.off()}
\end{Verbatim}
\begin{Verbatim}[commandchars=\\\{\}]
\textcolor{blue}{R> pdf(file="k7_6_n100.pdf") ; knotplot2(k7_6,n=100) ; dev.off()}
\end{Verbatim}

Figure 4 caption: Much of the explanation of ``7 3" versus ``3 7" is
already in the main text.  \textcolor{blue}{\\Fixed\\ \\}

Page 8, Line 60: ``Why 64 real variables"?  What are they?  How does
this agree with ``R50" on page 6?\textcolor{blue}{\\ The 64 variables
  are just the parameters required to describe the knot path, it
  agrees with the (reworded) coment about $\mathbb{R}^{50}$.\\ \\}

### 2023-02-06 inst/jma_knot.tex
pursuant to #15
typo

\begin{description}
\item[{\tt total\_crossing\_potential\_energy()}] The idea is to make
  the crossing points attract one another as if they are
  gravitationally active objects; metric measured by their mutual
  potential energy.  If crossing point $i$ is at position ${\mathbf
    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
    p}_j\right|^{2}$.  The idea is to encourage more compact
  configurations rather than long skinny ones.
\item[{\tt node\_crossing\_badness()}] Here we make the crossing
  points {\em repel} one another, as though they are all charged
  objects (with the same sign).  The idea is to spread out the nodes
  so no two nodes are too close together.  If crossing point $i$ is at
  position ${\mathbf p}_i$ we have $\sum_{i\neq j}\left|{\mathbf
    p}_i-{\mathbf p}_j\right|^{-1}$.  The length scale for this
  repulsion is smaller than that of the attractive term above.  The
  overall potential thus resembles that sometimes seen in atomic
  physics when considering the hydrogen molecule $H_2$: the two
  protons attract one another when distant, but repel when closer.  In
  the package, the intent is for this optimal distance to be a shallow
  minimum and the optimization routine is free to adjust within a wide
  range.
\item[{\tt total\_crossing\_angles()}] Here we wish to punish
  configurations with strands that cross at acute angles.  If crossing
  point $i$ has crossing angle $\theta_i$ the metric is
  $\sum_i\left(\cos\theta_i\right)^2$.
\item[{\tt total\_bending\_energy()}] The idea is for the strands to
  preferentially follow elastica curves, as in elementary (nonlinear)
  beam theory.  From differential geometry, the Euler elastica
  potential energy is $\int\kappa(s)^2\,ds$.  In the package, this is
  calculated numerically for each strand and the total returned.
\item[{\tt total\_string\_length()}] The length is just $\int ds$, with a
  nominal value of 5000 (which value is about right for a printed
  copy).
\item[{\tt midpoint\_badness()}] The idea is to suppress strands with
  crossing points too close to the endpoint nodes.  For each crossing
  point $c_{ij}$ of strand $i$ with strand $j$, we calculate the arc
  length $\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
  point and normalize with the strand length $\int\,ds$.  We reward
  configurations with normalized crossing point arc lengths that are
  close to $1/2$.  Note that this particular metric is not independent
  of the total crossing potential energy metric.
\item[{\tt curvature\_switching\_badness()}] Here the problem is knots
  which should have either all their strands curving to the left, or
  all to the right; the signed curvature should be everywhere the same
  sign (the canonical example would be $8_{18}$ which in earlier
  versions of the software had ugly changes of curvature).  We impose
  a fixed penalty for any knot with more than one sign.
\item[{\tt curvature\_consecutive\_segment\_switching\_badness()}] We
  wish to discourage configurations in which curvature undergoes a
  sudden change at a node.  The value for a single (Bezier) node is
  simply $\left(\delta\kappa\right)^2$, and the penalty for a given
  configuration is the sum of the values for each node.  See
  Figure~\ref{two_trefoil_knots} for motivation and discussion.
\item[{\tt non\_crossing\_strand\_close\_approach\_badness()}] Here
  the idea is to prevent non-intersecting strands from approaching one
  another too closely.  Quite apart from aesthetic considerations,
  this metric prevents the creation of new spurious nodes that would
  disrupt the optimization routines.
\end{description}

@@

>>>>

\begin{description}
\item[{\tt total\_crossing\_potential\_energy()}] The idea is to make
  the crossing points attract one another as if they are
  gravitationally active objects; metric measured by their mutual
  potential energy.  If crossing point $i$ is at position ${\mathbf
    p}_i$ we have $\sum_{i\neq j}\left|{\mathbf p}_i-{\mathbf
    p}_j\right|^{2}$.  The idea is to encourage more compact
  configurations rather than long skinny ones.
\item[{\tt node\_crossing\_badness()}] Here we make the crossing
  points {\em repel} one another, as though they are all charged
  objects (with the same sign).  The idea is to spread out the nodes
  so no two nodes are too close together.  If crossing point $i$ is at
  position ${\mathbf p}_i$ we have $\sum_{i\neq j}\left|{\mathbf
    p}_i-{\mathbf p}_j\right|^{-1}$.  The length scale for this
  repulsion is smaller than that of the attractive term above.  The
  overall potential thus resembles that sometimes seen in atomic
  physics when considering the hydrogen molecule $H_2$: the two
  protons attract one another when distant, but repel when closer.  In
  the package, the intent is for this optimal distance to be a shallow
  minimum and the optimization routine is free to adjust within a wide
  range.
\item[{\tt total\_crossing\_angles()}] Here we wish to punish
  configurations with strands that cross at acute angles.  If crossing
  point $i$ has crossing angle $\theta_i$ the metric is
  $\sum_i\left(\cos\theta_i\right)^2$.
\item[{\tt total\_bending\_energy()}] The idea is for the strands to
  preferentially follow elastica curves, as in elementary (nonlinear)
  beam theory.  From differential geometry, the Euler elastica
  potential energy is $\int\kappa(s)^2\,ds$.  In the package, this is
  calculated numerically for each strand and the total returned.
\item[{\tt total\_string\_length()}] The length is just $\int ds$, with a
  nominal value of 5000 (which is about right for a printed copy).
\item[{\tt midpoint\_badness()}] The idea is to suppress strands with
  crossing points too close to the endpoint nodes.  For each crossing
  point $c_{ij}$ of strand $i$ with strand $j$, we calculate the arc
  length $\int_{s=0}^{c_{ij}}\,ds$ from the node to each crossing
  point and normalize with the strand length $\int\,ds$.  We reward
  configurations with normalized crossing point arc lengths that are
  close to $1/2$.  Note that this particular metric is not independent
  of the total crossing potential energy metric.
\item[{\tt curvature\_switching\_badness()}] Here the problem is knots
  which should have either all their strands curving to the left, or
  all to the right; the signed curvature should be everywhere the same
  sign (the canonical example would be $8_{18}$ which in earlier
  versions of the software had ugly changes of curvature).  We impose
  a fixed penalty for any knot with more than one sign.
\item[{\tt curvature\_consecutive\_segment\_switching\_badness()}] We
  wish to discourage configurations in which curvature undergoes a
  sudden change at a node.  The value for a single (Bezier) node is
  simply $\left(\delta\kappa\right)^2$, and the penalty for a given
  configuration is the sum of the values for each node.  See
  Figure~\ref{two_trefoil_knots} for motivation and discussion.
\item[{\tt non\_crossing\_strand\_close\_approach\_badness()}] Here
  the idea is to prevent non-intersecting strands from approaching one
  another too closely.  Quite apart from aesthetic considerations,
  this metric prevents the creation of new spurious nodes that would
  disrupt the optimization routines.
\end{description}

@@

\section*{Acknowledgement}

I would like to thank an anonymous referee whose insightful comments
led to a much improved manuscript.
